### 2019-09-18 reports/thesis/chapters/4_variancePrediction.tex
Check algebra

The aim is to model and predict the grey value variance of a pixel given the pixel's grey value by using a model fitted onto the sample variance-mean data. No spatial information was used, thus, the variance-mean data is denoted as $(x_1,y_1),(x_2,y_2),\dotsc,(x_N,y_N)$ from now on, where $N$ is the area of the ROI or the size of the dataset. Let $Y(x)$ be a random variable and the sample variance given a grey value mean $x$. It was assumed that the standard error from estimating the mean was negligible so that the uncertainty is captured by the random variable $Y$. It was assumed that for a given pixel, the grey values are Normal and i.i.d. Let $\sigma^2(x)$ be the variance given a grey value $x$, then it can be shown that
\begin{equation}
\dfrac{(m-1)Y(x)}{\sigma^2(x)}\sim\chi^2_{m-1}
\end{equation}
which results in
\begin{equation}
Y(x)\sim\gammaDist\left(\alpha,\dfrac{\alpha}{\sigma^2(x)}\right)
\end{equation}
where $\alpha=(m-1)/2$ is the shape parameter. The expectation and variance are
\begin{equation}
\expectation\left[Y(x)\right]=\sigma^2(x)
\end{equation}
and
\begin{equation}
\variance\left[Y(x)\right]=\dfrac{\sigma^2(x)}{\alpha}
\end{equation}
respectively.

@@

An example of a systematic component are polynomial features $\eta(x)=\sum_{r=0}^{p}\beta_r x^{r}$ so that when used with the identity link, for example, $\expectation\left[Y(x)\right] = \sum_{r=0}^{p}\beta_r x^{r}$. Iterative reweighted least squares \citep{friedman2001elements} can be used to estimate the parameters $\beta_0, \beta_2, \dotsc, \beta_{p}$ given data for the model to fit onto.

@@

The AIC and BIC are given as
\begin{equation}
\AIC = 2k-2\ln L
\end{equation}
and
\begin{equation}
\BIC = k\ln m - 2\ln L
\end{equation}
respectively where $k$ is the number of parameters in the systematic component and $\ln L$ is the log-likelihood of the GLM. The model with the lowest AIC or BIC is preferred. GLM aims to maximise the log-likelihood but the additional terms in the criteria penalise models with too many terms. The log-likelihood is given as
\begin{equation}
  \ln L = \sum_{i=1}^N \left[
    \alpha\ln\alpha
    -\ln\Gamma(\alpha)
    -\alpha\ln \widehat{y}_i
    +(\alpha-1)\ln y_i
    -\frac{\alpha y_i}{\widehat{y}_i}
  \right]
  \ .
\end{equation}
where $\widehat{y}_i=\widehat{y}(x_i)$. $\alpha=(m-1)/2$ was assumed to be known so does not need to be estimated.

@@

The mean scaled deviance was used to assess the performance of variance prediction. The deviance is defined to be
\begin{equation}
  D = 2\left(
    \ln L_\mathrm{s} - \ln L    
  \right)
\end{equation}
where $L$ and $L_\mathrm{s}$ are the likelihood and saturated likelihood respectively.  The saturated likelihood is obtained by replacing all $\widehat{y}_i$ with $y_i$ in the likelihood so that
\begin{equation}
  \ln L_\mathrm{s} = \sum_{i=1}^N \left[
    \alpha\ln\alpha
    -\ln\Gamma(\alpha)
    -\alpha\ln{y_i}
    +(\alpha-1)\ln y_i
    -\alpha
  \right]
  \ .
\end{equation}
Following from this, the deviance is
\begin{equation}
  D = 2\alpha
  \sum_{i=1}^N\left[
      \dfrac{
          y_i-\widehat{y}_i
      }
      {
          \widehat{y}_i
      }
      - \ln\left(\dfrac{y_i}{\widehat{y}_i}\right)
  \right]
  \ .
\end{equation}
The mean scaled deviance is obtained by removing the factor of $\alpha$ and dividing by $N$ to get
\begin{equation}
    D_\mathrm{s} = \dfrac{2}{N}
    \sum_{i=1}^N\left[
        \dfrac{
            y_i-\widehat{y}_i
        }
        {
            \widehat{y}_i
        }
        - \ln\left(\dfrac{y_i}{\widehat{y}_i}\right)
    \right]
    \ .
\end{equation}

>>>>

The aim is to model and predict the grey value variance of a pixel given the pixel's grey value by using a model fitted onto the sample variance-mean data. No spatial information was used, thus, the variance-mean data is denoted as $(x_1,y_1),(x_2,y_2),\dotsc,(x_N,y_N)$ from now on, where $N$ is the area of the ROI or the size of the dataset. Let $Y(x)$ be a random variable and the sample variance given a grey value mean $x$. It was assumed that the standard error from estimating the mean was negligible so that the uncertainty is captured by the random variable $Y$. It was assumed that for a given pixel, the grey values are Normal and i.i.d. Let $\sigma^2(x)$ be the variance given a grey value $x$, then it can be shown that
\begin{equation}
\dfrac{(m-1)Y(x)}{\sigma^2(x)}\sim\chi^2_{m-1}
\end{equation}
which results in
\begin{equation}
Y(x)\sim\gammaDist\left(\alpha,\dfrac{\alpha}{\sigma^2(x)}\right)
\end{equation}
where $\alpha=(m-1)/2$ is the shape parameter. The expectation and variance are
\begin{equation}
\expectation\left[Y(x)\right]=\sigma^2(x)
\end{equation}
and
\begin{equation}
\variance\left[Y(x)\right]=\left(\dfrac{\sigma^2(x)}{\alpha}\right)^2
\end{equation}
respectively.

@@

An example of a systematic component are polynomial features $\eta(x)=\sum_{r=0}^{p}\beta_r x^{r}$ so that when used with the identity link, for example, $\expectation\left[Y(x)\right] = \sum_{r=0}^{p}\beta_r x^{r}$. Iterative reweighted least squares \citep{friedman2001elements} can be used to estimate the parameters $\beta_0, \beta_1, \dotsc, \beta_{p}$ given data for the model to fit onto.

@@

The AIC and BIC are given as
\begin{equation}
\AIC = 2k-2\ln L
\end{equation}
and
\begin{equation}
\BIC = k\ln N - 2\ln L
\end{equation}
respectively where $k$ is the number of parameters in the systematic component and $\ln L$ is the log-likelihood of the GLM. The model with the lowest AIC or BIC is preferred. GLM aims to maximise the log-likelihood but the additional terms in the criteria penalise models with too many terms. The log-likelihood is given as
\begin{equation}
  \ln L = \sum_{i=1}^N \left[
    \alpha\ln\alpha
    -\ln\Gamma(\alpha)
    -\alpha\ln \widehat{y}_i
    +(\alpha-1)\ln y_i
    -\frac{\alpha y_i}{\widehat{y}_i}
  \right]
  \ .
\end{equation}
where $\widehat{y}_i=\widehat{y}(x_i)$. $\alpha=(m-1)/2$ was assumed to be known so does not need to be estimated.

@@

The mean scaled deviance was used to assess the performance of variance prediction. The deviance is defined to be
\begin{equation}
  D = 2\left(
    \ln L_\mathrm{s} - \ln L
  \right)
\end{equation}
where $L$ and $L_\mathrm{s}$ are the likelihood and saturated likelihood respectively.  The saturated likelihood is obtained by replacing all $\widehat{y}_i$ with $y_i$ in the likelihood so that
\begin{equation}
  \ln L_\mathrm{s} = \sum_{i=1}^N \left[
    \alpha\ln\alpha
    -\ln\Gamma(\alpha)
    -\ln y_i
    -\alpha
  \right]
  \ .
\end{equation}
Following from this, the deviance is
\begin{equation}
  D = 2\alpha
  \sum_{i=1}^N\left[
      \dfrac{
          y_i-\widehat{y}_i
      }
      {
          \widehat{y}_i
      }
      - \ln\left(\dfrac{y_i}{\widehat{y}_i}\right)
  \right]
  \ .
\end{equation}
The mean scaled deviance is obtained by removing the factor of $\alpha$ and dividing by $N$ to get
\begin{equation}
    D_\mathrm{s} = \dfrac{2}{N}
    \sum_{i=1}^N\left[
        \dfrac{
            y_i-\widehat{y}_i
        }
        {
            \widehat{y}_i
        }
        - \ln\left(\dfrac{y_i}{\widehat{y}_i}\right)
    \right]
    \ .
\end{equation}

### 2019-09-18 reports/thesis/chapters/5_inference.tex
Check algebra

for $r=0,1,2,3,\dotsc$ until some convergence condition is met. The derivatives for the log density was obtained via the following. Recall that the kernel density estimate is $\widehat{p}_Z(z)=
  \frac{1}{nh}
  \sum_{i=1}^n\phi\left(
    \frac{z_i-z}{h}
  \right) $,
so the log density is

>>>>

for $r=0,1,2,3,\dotsc$ until some convergence condition is met. The derivatives for the log density were obtained via the following. Recall the kernel density estimate in Equation \eqref{eq:inference_kernel_density_estimate}, the log density is
